---
title: "Eelgrass ecological status"
subtitle: |
  [NO_EELG_001] 
# delete the line with the second indicatorID (line 5) 
# if it's not relevant for you.
# In the parentheses you can add a word to distinguish multiple indicators 
# from each other, for example the ecosystem type that they cover
indicatorID: 
  - NO_EELG_001
# If you are presenting multiple indicators, and have multiple 
# indicator IDs, then add them as a list here, and make sure you add
# the same number of list entries, 
# and in the same order that corresponds to the indicator IDs,  
# to the meta data fields where the indicators differ. 
# For example, if the indicators cover different ecosystem types, 
# add the ecosystems as a list under 'Ecosystems', but if they are in the
# same realm, same country etc., then you only need to write the realm and
# country once. 
indicatorName: 
  - Eelgrass ecological status # add more by adding another line starting '-'
verbatimName: 
  - Økologisk status av ålegress # the indicator name that us used in the native language
folderName: XXX_YYY_ZZZ # Leave to be filled out by ecRxiv during typesetting (name of the folder under 'indicators/'). 
# If it's a single indicator being reported, the folder name should be the same as the indicatorID. 
# If there are multiple ID's then the folder name could look something like XX_YYY_ZZZ-ÅÅÅ.
country: Norway
continent: Europe
ECT: B2 - Structural state characteristics # e.g. A2 - Chemical State characteristics
Realm: # Please copy and paste Realms, Biomes and Ecostsem names from here: https://github.com/NINAnor/ecRxiv/wiki#ecosystem-typology 
  - Marine (M) # e.g. Freshwater (F)
Biome: 
  - Marine shelf biome # e.g. F1 Rivers and streams biome
Ecosystem: 
  - M1.1 Seagrass meadows # e.g. F1.3 Freeze-thaw streams
verbatimEcosystem: Ålegresseng (Saltvanns-undervannseng in NiN) 
# add name of the ecosystem or naturetype according to any relevant local or national typology. E.g. fjell, skog, ... 
yearAdded: 2025 # the year when the indicator was first submitted to ecRxiv
yearLastUpdate: 2025
status: incomplete # options are complete, incomplete, or deprecated
Version: "0.1.0" # See wiki page: https://github.com/NINAnor/ecRxiv/wiki#versioning
VersionComment: Comment here # Describe briefly how this version differs from earlier versions 
url: https://github.com/NINAnor/ecRxiv/tree/main/indicators/XX_YYYY_ZZZ
# To be filled out by ecRxiv during the typesetting, when they will replace XX_YYYY_ZZ with the name of the root folder for the current indicator(s). It should be the same as the indicatorID, but if there are more indicator IDs presented in the same folder, the folder name might be different, e.g. XX_YYYY_AAA-BBB 
Normalised: Yes # Yes or no depending on whether the metric has been normliased or not.
SpatialAggregationPathway: Transform -Scale - Aggregate # Describe the aggregation using the standrdaise terminology 
# e.g. Scale - Transform - Aggregate (length-weighted arithmetic mean) - Truncate
Precision: ???? #Describe the precition metric you have used. Precision may be reported either as variation among observed values (e.g. standard deviation, variance, range, interquartile range) or as uncertainty around an estimated value (e.g. standard error, confidence interval, credible interval). 
format: 
  html:
    embed-resources: true
    code-fold: true
    toc: true
    toc-title: Contents
    toc-depth: 3
    smooth-scroll: true
execute: 
  cache: true
author:
  - name: Dag Hjermann              # Enter name
    email: dag.hjermann@niva.no  # Enter email
    affiliations:
      - id: 1
        name: Norwegian Institute for Water Research # Enter affiliations
AuthorList: Hjermann, D.H. og ????  
date: September 17, 2026   # January 1, 1901 # Enter date
data_availability: none
code_reproducibility: none
open_science_badge: none
callout-icon: false
lightbox: true
css: ../../../style.css
code-links:
      - text: Add a review
        icon: github
        href: https://github.com/NINAnor/ecRxiv
hide: TRUE # Change to FALSE
bibliography: references.bib
codeReviewers:
  - name: Jane Smith
    email: jane.smith@example.no
    affiliations:
      - id: reviewerInst
        name: Example Institute
    reviewType: code review
    github: janesmith
  - name: John Doe
    affiliations:
      - ref: myID
    reviewType: reproducibility check
    github: johndoe
---

<!--# This is a template for how to document the indicator analyses. Make sure also to not change the order, or modify, the headers, unless you really need to. This is because it easier to read if all the indicators are presented using the same layout. If there is one header where you don't have anything to write, just leave the header as is, and don't write anything below it. If you are providing code, be careful to annotate and comment on every step in the analysis. Before starting it is recommended to fill in as much as you can in the metadata file. This file will populate the initial table in your output.-->

<!--# Load all you dependencies here -->

```{r setup}
#| include: false
#| cache: false
library(knitr)
library(tidyverse)
library(kableExtra)
library(here)
library(anybadger)
library(yaml)
library(tibble)
library(conflicted)
# weird workarounds for cache issues
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::lag)
conflicts_prefer(dplyr::select)
pull<-dplyr::pull

knitr::opts_chunk$set(echo = TRUE)

source("functions.R")    # self-defined functions

```

<!--# The following parts are autogenerated. Do not edit. -->


```{r source}
#| echo: false
#| cache: false
source(here::here("_common.R"))
```

```{r reviewers}
#| echo: false
#| results: asis

if (!is.null(yaml_data$codeReviewers)) {
  reviewers <- purrr::map_chr(yaml_data$codeReviewers, "name")
  
  htmltools::HTML(paste0(
    '<div id="title-block-header" class="quarto-title-block default">
       <div class="quarto-title-meta">
         <div>
           <div class="quarto-title-meta-heading">Reviewed by</div>
           <div class="quarto-title-meta-contents">',
    paste(reviewers, collapse = "; "),
    '</div>
         </div>
       </div>
     </div>',
    sep = ""
  ))
}
```

::: {layout-ncol="6"}

```{r statusBadge}
#| echo: false
status_badge(as.character(st))
```

```{r versionBadge}
#| echo: false
version_badge(my_version_number = as.character(version[[1]]))
```

```{r dataBadge}
#| echo: false
data_badge(as.character(badges[[1]]))
```

```{r codeBadge}
#| echo: false
code_badge(as.character(badges[[2]]))
```

```{r openScienceBadge}
#| echo: false
open_science_badge(as.character(badges[[3]]))
```

```{r licenseBadge}
#| echo: false
# The following inserts a badge specifying the GPL GNU version 3 license. ecRxiv recommends using this license, but you can decide to use another. Contact ecRxiv for help in insert a badge for a different license
license_badge()
```


::: 
 

> **Recommended citation**: `r paste(auth, " ", year, ". ", name, " (ID(s): ", folder_name, ") ", "v. ", version, ". ecRxiv: ", url, sep="")`



::: {.callout-tip collapse="true"}
## Metadata
```{r tbl-meta2}
#| tbl-cap: 'Indicator metadata'
#| echo: false
#| warning: false
meta |>
  dplyr::filter(Variable %in% c(
    "Indicator ID",
    "Indicator Name",
    "verbatimName",
    "Continent",
    "Country",
    "Ecosystem Condition Typology Class", 
    "Realm",
    "Biome",
    "Ecosystem",
    "verbatimEcosystem",
    "Year added", 
    "Last update",
    "Version",
    "Version comment",
    "Normalised",
    "Spatial aggregation pathway",
    "Precision")) |>   
  kable()

```
:::

::: {.callout-tip collapse="true"}

## Log

<!--# Update this logg with short messages for each update -->
- 01 Jan. 1901 - Original PR
:::


<hr />

<!--# Document you work below.  -->

## 1. Summary

<!--# 
With a maximum of 300 words, describe the indicator in general terms as you would to a non-expert. Think of this as a kind of common language summary. Follow the four numbers sub-headers below. under point 2, it is possible under  include a bullet point list of the specific steps in the workflow. 
 -->

1. **Background and rationale**:
The eelgrass index (Ålegressindeks) is a biological indicator used to assess ecological condition in Norwegian coastal waters under the Water Framework Directive and the Norwegian Water Regulation. The indicator is based on eelgrass (Zostera marina), a habitat-forming marine angiosperm that occurs on soft bottom in sheltered coastal and fjord environments where macroalgal indicators are often absent. Eelgrass meadows are important habitats for many marine animals including juvenile fish. They are also sensitive to eutrophication, reduced water transparency and increased growth of filamentous algae, making them effective indicators of environmental degradation.  

2. **Methods**:
The indicator combines three components of eelgrass condition: lower depth limit, meadow density and degree of epiphytic algal overgrowth. Lower depth limit reflects water clarity and light availability, density reflects meadow integrity, and overgrowth reflects eutrophication-related stress. Each component is converted to a score based on reference conditions and class boundaries defined for Norwegian coastal ecoregions and water types. The final Ecological Quality Ratio (EQR) is calculated as a weighted index where depth limit contributes 50%, density 30% and overgrowth 20%. EQR values range from 0 to 1, where higher values indicate better ecological condition.


3. **Interpretation**:
<!-- Add a very short description of how the indicator values should be interpreted. Here's an example for a bumblebee indicator: An indicator value of 1 signifies an intact bumblebee community. Values decreasing toward zero reflect negative ecological change, specifically the decline of species expected to be present. This metric does not account for potentially positive changes, such as increases in species' abundance or the arrival of new species. Tip: Add text on the line directly following the subheader, and don't introduce blank lines. This way the whole paragraph renders as a block quote. -->
An EQR value close to 1 indicates eelgrass meadows with near-reference depth distribution, high density and little algal overgrowth. Lower values indicate increasing deviation from reference conditions, typically associated with eutrophication, reduced light penetration and habitat degradation.  

4. **Results and development needs**:
<!-- Briefly, what does the indicator say about the ecosystem condition. What is the current status of  the metric (can it be used or is it still in development), and what immediate tasks remain to improve the indicator or to make the indicator operational -->
The current implementation uses Vannmiljø monitoring data from the Økokyst (and Oslofjord??) reference stations and provides a transparent and reproducible workflow for calculating eelgrass EQR values. Future development should focus on broader geographic coverage, automated linkage to national typology datasets and formal uncertainty estimation.

## 2. About the underlying data

<!--# Give a general introduction the datasets you have used, and provide more specific details in the sub-chapters.  -->

Data are obtained from Vannmiljø observations originating primarily from Økokyst monitoring activities. Observations are recorded at individual monitoring stations and sampling dates. The current implementation includes repeated measurements from multiple years at fixed eelgrass locations.

### 2.1 Spatial and temporal resolution and extent

<!--# Describe the temporal and spatial resolution and extent of the data used -->

Eelgrass meadows occur as distinct patches in the marine environment. Measurements of ecological status is collected at a limited number of selected eelgrass meadows, here denoted as monitoring stations. The monitoring stations are fixed and data for the stations are collected once every year or less frequently.  


### 2.2 Sampling design

<!-- How was the data sampled? Describe potential sampling biases -->

#### Measurements in the field  

Surveys are conducted from small boats using underwater video equipment with an integrated depth sensor. Monitoring follows protocols developed for Norwegian coastal ecosystem monitoring and the Water Framework Directive. The observations used by the indicator are collected along a set of standardized transects designed to capture variation in the lower depth limit of eelgrass meadows, eelgrass density, and algal overgrowth throughout the meadow.

At each monitoring station, observations are made along one transect following the shoreline and an additional 5–10 transects extending from shallow water to slightly below the lower depth limit of the eelgrass meadow. Along these transects, point observations are recorded using an underwater camera, while all positions are georeferenced using GPS. The approach is intended to provide representative coverage of the entire eelgrass meadow, including both the shallow portions and the deepest occurrence of eelgrass.

The lower depth limit used in the indicator corresponds to the lower depth limit of the eelgrass meadow rather than the deepest individual plant. Following the national methodology, the lower meadow limit is defined as the deepest observation classified as sparse eelgrass (density class 2). This definition is intended to provide a robust and repeatable estimate of the depth distribution of the meadow.

Eelgrass density and epiphytic algal overgrowth are recorded at all observation points along the shoreline transect and the offshore transects. Both variables are classified using a four-level ordinal scale similar to that employed in the Norwegian National Programme for Mapping Biodiversity in Coastal Waters. Eelgrass density ranges from isolated plants to dense, continuous meadows, while overgrowth ranges from little or no epiphyte cover to dominant algal overgrowth.

To ensure comparability among stations and years, all depth measurements are corrected for water level variation using information on tides, wind conditions, and atmospheric pressure. This standardization reduces variation caused by short-term changes in sea level and allows ecological changes in eelgrass depth distribution to be assessed more consistently over time.

Detailed descriptions of field procedures, point selection criteria, and calculation rules are provided in Trannum et al. (2022).


### 2.3 Original units

<!--# What are the original units for the most relevant  variables in the data-->

**The lower depth limit** is recorded in meters (m).  

**Eelgrass density** is measeured on a 4 level scale:  

| Level | Description |
|---------|-------------|
| 1 | Enkeltfunn (isolated plants) |
| 2 | Spredte planter (sparse meadow) |
| 3 | Flekkvis tett eng (patchy dense meadow) |
| 4 | Tett, heldekkende eng (continuous dense meadow) |

**Epiphytic algal overgrowth** is measured on a 4 level scale:

| Level | Observed category | Observed category (No) | Indicator score |
|------------------|----------------|----------------|
| 0 | No overgrowth | Ingen begroing | 4 |
| 1 | Little overgrowth | Lite begroing | 4 |
| 2 | Sparse overgrowth |  Spredt begroing | 3 |
| 3 | Common overgrowth | Vanlig begroing | 2 |
| 4 | Dominating overgrowth | Dominerende begroing | 1 |


### 2.4 Instructions for citing, using and accessing data

<!--# Is the data openly available? If not, how can one access it? What are the key references to the datasets?   -->

The data can be downloaded from The Environment Agency's database *Vannmiljø* at https://vannmiljo.miljodirektoratet.no/. They are licenced by Norsk lisens for offentlige data (NLOD) 2.0 (https://data.norge.no/nlod). 


### 2.5 Additional comments about the dataset

<!--# Optional. Text here -->

## 3. Indicator properties

### 3.1 Ecosystem Condition Typology Class (ECT)

<!--# Describe the rationale for assigning the indicator to the ECT class. See https://oneecosystem.pensoft.net/article/58218/
This doesnt need to be very long. Maybe just a single sentence. -->

Seagrass meadows are classified as an ecosystem in the Global Ecosystem Typology. In Norway, eelgrass meadows is the main type of seagrass meadows and the seagrass is the main structural element of that habitat.

### 3.1.1 Other condition typologies

<!--# Optional: Add text about other standards or typologies, such as alternative national condition typologies, and how the indicator relates to these -->


### 3.2 Ecosystem condition characteristic

<!--# 

A short (one to three word) tag or name for the ecosystem condition characteristic represented in the indicator. See 10.3897/oneeco.6.e58218 for information on what these characteristics might be.

Examples: 'intact hydrology'; 'near-natural species composition'; 'soil stability'; sustainable trophic interactions'. 

The term 'characteristic' is used in a similar way  to the term 'criteria' in Multiple Criteria Decision Making.  

-->

### 3.3 Collinearities with other indicators

<!--# Describe known collinearities with other metrics (indicators or variables) that could become problematic if they were also included in the same Ecosystem Condition Assessment as the indicator described here. -->


### 3.4 Impact factors

<!--# Describe the main natural and anthropogenic factors that affecst the metric -->

### 3.5 Spatial aggregation

<!-- Describe in word the same aggregation pathway presented in the top yaml (the metadata). If you have scaled against the reference value, it is especially important to know if you did any avergaing or summations etc. beforehand. An example text for a rather simple case could be: "The variable is scaled at the initial polygon level and the point estimate is a area-weighted arithmetic mean of all polygins. There is no transformation, and truncation is not needed as the indicator becomes naturally bound between 0 and 1".  -->  

## 4. Reference condition and levels

### 4.1 Reference condition

<!--# Define the reference condition (or refer to where it is defined). Note the distinction between reference condition and reference levels 10.3897/oneeco.5.e58216  -->

?????

### 4.2 Reference levels

<!--# 

If relevant (i.e. if you have normalised a variable), describe the reference levels used to normalise the variable. 

Use the terminology where X~0~ refers to the reference level (the variable value) denoting the worst possible condition; X~100~denotes the optimum or best possible condition; and X~*n*~, where in is between 0 and 100, denotes any other anchoring points linking the variable scale to the indicator scale (e.g. the threshold value between good and bad condition X~60^). 

Why was the current option chosen and how were the reference levels quantified? If the reference values are calculated as part of the analyses further down, please repeat the main information here.

-->

**Note terminology `X~0~`**  - TODO  

The reference levels for the lower depth limit of the eelgrass meadow (i.e. the depth limit for under pristine conditions) depends on ecoregion and water type, as the light penetration varies under natural conditions.  


#### Calculation of nEQR  

The final Ecological Quality Ratio (EQR) is calculated as a weighted combination of the three component scores:

$$ nEQR = 0.5 \cdot D +  0.3 \cdot T +  0.5 \cdot B$$ {#eq-continuity}

where:

* *D* = depth-limit score (0–5),
* *T* = density score (1–4),
* *B* = overgrowth score (1–4).

The weights reflect the relative importance assigned in the national classification scheme, with lower depth limit contributing 50% of the final score, density 30%, and algal overgrowth 20%. The resulting EQR is bounded between 0 and 1 and provides a normalized measure of ecological condition. Higher values indicate eelgrass meadows that more closely resemble reference conditions.


The **depth limit score (D)** depends on eco-region and water type:

| Ecoregion | Water type | Score 5 | Score 4 | Score 3 | Score 2 | Score 1 | Score 0 |
|-----------|----------|----------|----------|----------|----------|----------|----------|
| Nordsjøen sør | 2 | > 8 m | > 6–8 m | > 5–6 m | > 3–5 m | > 2–3 m | 0–2 m |
| Nordsjøen sør | 3 | > 8 m | > 6–8 m | > 5–6 m | > 3–5 m | > 2–3 m | 0–2 m |
| Nordsjøen nord | 3 | > 7 m | > 6–7 m | > 4–6 m | > 3–4 m | > 1–3 m | 0–1 m |
| Norskehavet sør* | 3 | > 7 m | > 6–7 m | > 4–6 m | > 3–4 m | > 1–3 m | 0–1 m |

The **density score (T)** is equal to the density, measured on a scale from 1-4. 

The **overgrowth score (B)** is given as follows::

| Level | Observed category | Overgrowth score (B) |
|------------------|----------------|----------------|
| 0 | No overgrowth | 4 |
| 1 | Little overgrowth | 4 |
| 2 | Sparse overgrowth |  3 |
| 3 | Common overgrowth | 2 |
| 4 | Dominating overgrowth | 1 |





#### 4.2.1 Spatial resolution and validity

<!--# 

Describe the spatial resolution of the reference levels. E.g. is it defined as a fixed value for all areas, or does it vary. Also, at what spatial scale are the reference levels valid? For example, if the reference levels have a regional resolution (varies between regions), it might mean that it is only valid and correct to use for normalising local variable values that are first aggregated to regional scale. However, sometimes the reference levels are insensitive to this and can be used to scale variables at the local (e.g. plot) scale. 
 -->  

## 5. Uncertainties
### 5.1 General uncertainties
<!--# Describe the main reasons why you might not trust 100% the final value to represent the given ecosystem characteristic for the give ecosystem type. What aspects of this condition metric are most sensitive to subjective decitions? Is the spatial sampling sufficient? Does the choice of modeling framework affect the values? Is the metric relevant do describe ecosystem condition for the give ecosystem type?  -->

### 5.2 Reported uncertainties
<!-- Describe the nature of the errors you (hopefully) provide alongside the point estimate for the condition metric. It is especially important to clarify if the errors are inferential (referring to the uncertainty of the point estimate) or descriptive (referreing to the data itself). Here' an example text: "What we report as uncertainties in @tbl-finalTable are inferential and refer to the uncertainty around the point estimates. It does not directly describe variation in the data itself. 
However, the variation in the mean is defined by the spatial variation only, and quartiles are obtained by bootstrapping the set of localities within a region, and calculating the indicator for each of 9999 such samples".   -->



## 6. References

<!--# You can add references manually or use a citation manager and add intext citations as with crossreferencing and hyperlinks. See https://quarto.org/docs/authoring/footnotes-and-citations.html -->

## 7. Import and prepare datasets

<!--# All the data import and preparation should be done in this chapter, before the analysis chapter. As a minimum, add code for importing data, and describe or show the data structure. You may also explain certain variables, what they represent, and their distributions -->


### 7.1 Data from Vannmiljø  

<!--# Import and prepare the first dataset for analysis. Change the header from Dataset A to the name of the actual dataset. -->

```{r}

# read data in Vannmiljø format  

eelgrass_raw <-
  read_csv("../data/VM_export_eelgrass_station-based-search_2026-09-13.csv", 
    col_types = vm_file_specifications) %>%
  mutate(
    year = lubridate::year(Tid_provetak))

```


### 7.2. Station metadata (ecoregion and water type)

```{r}

# . # . # . # . # . # . # . # . # . # . # . # . # . # . # . # . # . 
# Station metadata ----
# . # . # . # . # . # . # . # . # . # . # . # . # . # . # . # . # . 

station_lookup <-
  tribble(
    ~Vannlokalitetsnavn, ~stasjonskode, ~okoregion, ~vanntype,
    "Midttunvågen",      "ZT34", "Nordsjøen nord", 3,
    "Nedstrand",         "ZT38", "Nordsjøen sør",  2,
    "Torsteinvika",      "ZT39", "Nordsjøen sør",  3,
    "Vikevågen",         "ZT37", "Nordsjøen sør",  2,
    "Jonebukta",         "ZT36", "Nordsjøen sør",  2,
    "Sundsbukta",        "ZR9",  "Norskehavet sør", 3,
    "Leksvik",           "ZT29", "Norskehavet sør", 3,
    "Sørhamna",          "ZR10", "Norskehavet sør", 3,
    "Storsletta",        "ZR11", "Norskehavet nord", 3,
    "Skjæret",           "ZR12", "Norskehavet nord", 3
  )


eelgrass_raw |>
  distinct(Vannlokalitetsnavn, year) |> 
  count(Vannlokalitetsnavn, name = "Number of years") |> 
  left_join(station_lookup, by = join_by(Vannlokalitetsnavn)) |> 
  filter(!is.na(okoregion))

```


## 8. Spatial units

<!--# Describe the spatial units that you rely on in your analyses. Highlight the spatial units (the resolution) that the indicator values should be interpreted at. Potential spatial delineation data (maksing layers etc.) should be introduced under chapert 7. We recommend using the SEEA EA terminology of Basic Spatial Units (BSU), Ecosystem Asses (EA) and Ecosystem Accounting Area (EAA). -->




## 9. Analyses

<!--# 

Use this header for documenting the analyses. Put code in separate code chunks, and annotate the code in between using normal text (i.e. between the chunks, and try to avoid too many hashed out comments inside the code chunks). Add subheaders as needed. 

Code folding is activated, meaning the code will be hidden by default in the html (one can click to expand it).

Caching is also activated (from the top YAML), meaning that rendering to html will be quicker the second time you do it. This will create a folder inside you project folder (called INDICATORID_cache). Sometimes caching created problems because some operations are not rerun when they should be rerun. Try deleting the cash folder and try again.

-->

```{r convert_to_wide}

# - convert 'long format' to 'wide' (one observation per station/year)

eelgrass_wide_1 <-
  eelgrass_raw %>%
  filter(
    Parameter_navn %in% c(
      "Absolutt nedre voksegrense angiospermer (takson)",
      "Nedre voksegrense ålegresseng (takson)",
      "Dekningsgrad ålegresseng (skala)",
      "Dekningsgrad marine begroingsalger (skala)"
    )
  ) %>%
  select(
    Vannlokalitetsnavn,
    Tid_provetak,
    year,
    Parameter_navn,
    Verdi
  ) %>%
  pivot_wider(
    names_from = Parameter_navn,
    values_from = Verdi
  )

```

```{r rename}

# Rename variables

eelgrass_wide_1 <- eelgrass_wide_1 |>
  rename(
    depth_limit_abs =
      `Absolutt nedre voksegrense angiospermer (takson)`,

    depth_limit =
      `Nedre voksegrense ålegresseng (takson)`,

    density =
      `Dekningsgrad ålegresseng (skala)`,

    epiphytes =
      `Dekningsgrad marine begroingsalger (skala)`
  )

# eelgrass_wide_1 |> head(3)


```

```{r add_station_metadata}

# add station metadata

eelgrass_wide_2 <- eelgrass_wide_1 |>
  left_join(
    station_lookup,
    by = "Vannlokalitetsnavn"
  ) %>%
  # remove stations without metadata
  # (these stations are not assessed in the Økokyst reports)
  filter(
    !is.na(okoregion)
  )

# eelgrass_wide_2 |> head(3)

```

The calculation of nEQR is based on `Nedre voksegrense ålegresseng (takson)`, `Dekningsgrad ålegresseng (skala)`, and `Dekningsgrad marine begroingsalger (skala)`, as given in 
```{r calculate_nEQR}

# calculate nEQR for each station / year  
# 
eelgrass_eqr <- eelgrass_wide_2 %>%
  mutate(

    depth_score =
      get_depth_score(
        depth_limit,
        okoregion,
        vanntype
      ),

    density_score =
      get_density_score(density),

    epiphyte_score =
      get_epiphyte_score(epiphytes),

    nEQR =
      0.5 * (depth_score / 5) +
      0.3 * (density_score / 4) +
      0.2 * (epiphyte_score / 4),
    
    # for ordering in tables
    okoregion = factor(okoregion,
                       levels = c("Norskehavet sør", "Nordsjøen nord", "Nordsjøen sør"))
    
  )

# head(eelgrass_eqr, 2)

```

As an example, we show the measured values (depth_limit, density, epiphytes), the scores (depth_score, density_score, epiphyte_score), and the nEQR for three stations.  
```{r results_by_station_2025}

# results_by_station, 2025

eqr_table1 <- eelgrass_eqr |>
  filter(!is.na(nEQR) & year == 2025) |>
  select(Vannlokalitetsnavn, okoregion,
         depth_limit, density, epiphytes,
         depth_score, density_score, epiphyte_score, nEQR) |>
  arrange(okoregion)

head(eqr_table1, 3) |> 
  kable()

```

```{r results_by_station_allyears }
# 
# results by station, all years

eqr_table2 <- eelgrass_eqr |>
  select(Vannlokalitetsnavn, okoregion, year, nEQR) |>
  filter(!is.na(nEQR)) |> 
  pivot_wider(names_from = year, values_from = nEQR) |>
  arrange(okoregion, Vannlokalitetsnavn) 

# eqr_table2

```

Below, we show the mean score for each of the three criteria for each ecoregion as well as the mean of the resulting nEQR.
```{r results_by_ecoregion_2025}

# results_by_station, 2025

eqr_region <- eelgrass_eqr |>
  filter(!is.na(nEQR) & year == 2025) |>
  group_by(okoregion) |> 
  summarise(
    depth_score = mean(depth_score, na.rm = TRUE), 
    density_score = mean(density_score, na.rm = TRUE), 
    epiphyte_score = mean(epiphyte_score, na.rm = TRUE),
    nEQR = mean(nEQR, na.rm = TRUE)
  ) |>
  arrange(okoregion) 

eqr_region |> 
  kable()

```

## 10. Results

<!--# 

Repeat the final results here. Typically this is a map or table of indicator values.

This is typically where people will harvest data from, so make sure to include all relevant output here, but don't clutter this section with too much output either.

-->

The results by ecoregion for 2025 are shown below.

```{r}

eqr_region |> 
  select(okoregion, nEQR) |> 
  kable()  

```


## 11. Export file

<!--# 

Optional: Display the code (don't execute it) or the workflow for exporting the indicator values to file. Ideally the indicator values are exported as a georeferenced shape or raster file with indicators values, reference values and errors. You can also chose to export the raw (un-normalised or unscaled variable) as a separate product. You should not save large sptaial output data on GitHub. You can use eval=FALSE to avoid code from being executed (example below - delete if not relevant) 

-->



```{r export}
#| eval: false
```


## 12. Author contributions

<!-- Fill in CReDIT and MeRIT statments to clarify the contributions from individual authors. For MeRIT, see https://www.nature.com/articles/s41467-023-37039-1. For CRedit, see https://credit.niso.org/. You can start with the template structur below, and just delete the rows you dont use.  -->

**CRediT statement**

* Conceptualization: [Initials]
* Methodology: [Initials]
* Data curation: [Initials]
* Formal analysis: [Initials]
* Investigation: [Initials]
* Software: [Initials]
* Validation: [Initials]
* Visualization: [Initials]
* Resources: [Initials]
* Writing – original draft: [Initials]
* Writing – review & editing: [Initials]
* Supervision: [Initials]
* Project administration: [Initials]
* Funding acquisition: [Initials]


**MeRIT attribution**

This part supplements the CRediT statement with more granularity for method section.

**Study design**
The study was designed by [AB, CD].

**Data collection / fieldwork**
Field data were collected by [AB, EF] between [years]. Sampling protocol was developed by [CD].

**Laboratory procedures**
DNA extraction was performed by [GH]. Sequencing was conducted by [IJ].

**Data processing**
Raw data were curated and cleaned by [EF]. Code for preprocessing was written by [KL].

**Statistical analysis**
All analyses were conducted by [CD] using [software], with verification by [AB].

**Visualization**
Figures were produced by [KL], with input from [CD].

**Reproducibility**
All code was reviewed and validated by [AB, CD].



::: {.callout-tip collapse="true"}
## Session Info
<!-- You can leave this last part as it is, unless you have a good reason to alter it. -->

This information is mainly aimed at developers that need to recreate the R environment that produced the html.
Important note: this information is about the R session that was running when the html was rendered, and if some of the code chunks were not evaluated when the html was rendered, they may have been executed in a different environment.  

```{r sessionInfo}
#| echo: false
sessionInfo()
```

:::
